---
title: "Getting started with stream"
author: "Michael Hahsler"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Getting started with stream}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(stream)
set.seed(1234)
```

The `stream` package provides tools to generate or read data streams, update
stream mining models, and inspect their results. This guide walks through a
small clustering workflow: create a stream, inspect incoming points, build a
clustering model, and evaluate it on new points.

## Installation

Install the released package from CRAN, then load it in your R session:

```{r install, eval=FALSE}
install.packages("stream")
```

```{r load-package}
library(stream)
```

## Create and inspect a data stream

`DSD_Gaussians()` generates a stream with three clusters and a small amount of
noise. A stream can be unbounded, so request a finite number of points with
`get_points()`:

```{r create-stream}
stream <- DSD_Gaussians(k = 3, d = 2, noise = 0.05)
points <- get_points(stream, n = 5)
points
```

The returned data frame contains feature columns and information columns such
as `.class`, which records the generating cluster. Use `info = FALSE` when you
only need feature values. Plot a sample of the stream to see its structure:

```{r plot-stream, fig.width=6, fig.height=5}
plot(stream, n = 500)
```

## Update a clustering model

`DSC_TwoStage()` combines an online micro-clusterer with an offline
macro-clusterer. Here, a sliding window stores recent points and k-means groups
the window's micro-clusters into three macro-clusters. `update()` feeds 500
points from the stream into the model:

```{r cluster-stream}
clustering <- DSC_TwoStage(
  micro = DSC_Window(horizon = 100),
  macro = DSC_Kmeans(k = 3)
)

update(clustering, stream, n = 500)
clustering
```

Inspect the number of clusters and their centers, then plot the result:

```{r inspect-clustering, fig.width=6, fig.height=5}
nclusters(clustering, type = "micro")
nclusters(clustering, type = "macro")
get_centers(clustering, type = "macro")
plot(clustering, stream, type = "both")
```

## Evaluate on new points

`evaluate_static()` tests the current model on new points without updating it.
The generated stream includes true cluster labels, so external measures such as
purity and the adjusted Rand index can be calculated:

```{r evaluate-clustering}
evaluate_static(
  clustering,
  stream,
  measure = c("purity", "crand"),
  n = 200,
  type = "macro",
  assign = "macro"
)
```

## Further reading

See the [function reference](../reference/index.html) and the package vignettes
for details on data generators, stream filters, clustering methods, evaluation,
and extending `stream` with new components.
